---
title: "German Willkommenskultur"
author: "Christopher Cosler and Lisa Schmid"
date: "December 4, 2015"
output:
  revealjs::revealjs_presentation:
    theme: sky
    transition: fade
    highlight: pygments
    center: true
---

```{r, include = FALSE}
# Load source file
source("Analysis/Analysis.R")
# Load libraries for presentation
library(stargazer)
```

# Introduction

## Level two if wanted

- Turn off alarm
- Get out of bed

## Level three if wanted

- Eat eggs
- Drink coffee

# Introduce the topic but make it interesting! (Lisa)

# Theory & Hypothesis (Lisa)

# Data sources (Chris)
We retrieve our data from three different sources:

1. *ProAsyl map* ^[https://www.google.com/maps/d/viewer?mid=zc6TdvfelKuY.kUvriXoSREXw]
2. *German Federal Agency for Cartography and Geodesy* ^[http://www.geodatenzentrum.de/geodaten/gdz_rahmen.gdz_div?gdz_spr=deu&gdz_akt_zeile=5&gdz_anz_zeile=1&gdz_unt_zeile=19&gdz_user_id=0]
3. *German Federal Statistical Office* ^[https://www-genesis.destatis.de/genesis/online]
4. *German Regional Statistical Offices* ^[https://www.regionalstatistik.de/genesis/online]

# Descriptive (Lisa)
## Map of Initatives
## Distribution of initatives per district (Hist)

#Inferential (Chris)

#Regression and why count
Test Test
```{r, echo=FALSE, results='asis'}
labels <- c("Old Age Dependency Ratio",
            'Gender Ratio',
            "Population with Abitur (percentage)",
            "GDP per Capita in 1000 Euros",
            'Unemployment Rate',
            'Population Density',
            'Number of Refugee Initiatives per District',
            "East-West Dummy",
            "Voter Turnout (percentage)")
stargazer(pois, qpois, nbin, 
          type='latex', # For PDF change style to latex, otherwise html
          title = "Poisson, Quasi-Poisson and Negative Binomial Regression Estimates for the number of Refugee Initiatives",
          digits = 2,
          header = FALSE,
          covariate.labels = labels,
          column.labels = c('Poisson', 'Quasi-Poisson', 'Negative Binomial'),
          model.names = FALSE,
          dep.var.labels = "Number of refugee initiatives per district"
          )
```
# Turnout rate
```{r, echo=FALSE}
plot_turnout
```

# East-West 

# 
#Descriptive (Chris)
##Maps

#Conclusion (Lisa)

#Shortcomings?

#Appendix
#Summary Table (Lisa)
